\documentclass[letterpaper, 11pt]{article}

\usepackage{amsmath}
\def\UrlBreaks{\do\/\do-}

%\usepackage[hyphens]{url}

\usepackage[margin=1in]{geometry}

%opening
\title{EEC134A Lab Report Guidelines}
\author{Instructor: Xiaoguang ``Leo'' Liu}
\date{Sept.~4, 2015}

\begin{document}

\maketitle

Lab reports are to be completed by teams not by individual students. 

You may follow the structure below for your lab report. 

\begin{enumerate}
	\item Cover
		\begin{enumerate}
			\item Include title, name of the team, name of the team members, and date 
		\end{enumerate}
	\item Procedures
		\begin{enumerate}
			\item There is no need to repeat the lab manual. If you followed the lab manual precisely, you can simply reference it, e.g. “Follow steps 1-5 of Experiment 4.2, …” 
			
			\item If you deviated from the lab manual, describe your procedures clearly and concisely. It may be a good idea to use photos to illustrate your procedures. You should also provide justifications why you followed different procedures. 
		\end{enumerate}
	\item Results and Discussions
		\begin{enumerate}
			\item Present your results in forms of tables, pictures, screen captures, data plots, etc. Pay attention to resolution, font size, trace width/color, etc to make sure that the plots are legible. All pictures must have a figure number and a caption. All pictures must be reference in the text.
			
			\item Discuss your results thoroughly. Some good questions to ask yourself while doing data analysis:
				\begin{enumerate}
					\item Are the results complete?
					\item Do the results make sense, i.e. do the results agree with theory/design? If yes, why? If no, why not?
					\item What are the trade-offs of a particular design? i.e. while you gained performance in some aspects, what price do you pay? 
				\end{enumerate}		
		\end{enumerate}
	\item References
		\begin{enumerate}
			\item Acknowledge everyone from whom you have benefited from in the completion of the lab report
			\item Reference all materials that are not generated by your team. 
		\end{enumerate}
	\item Appendix
		\begin{enumerate}
			\item Put design files and code in the appendix. 
		\end{enumerate}
\end{enumerate}

\end{document}
