\documentclass[letterpaper, 11pt]{article}

\usepackage{hyperref, xcolor, enumitem}
\usepackage[margin=1in]{geometry}

%opening
\title{EEC134 PCB Review Report Template}
\author{Instructor: Xiaoguang ``Leo'' Liu \\ lxgliu@ucdavis.edu \\ \small \href{http://creativecommons.org/licenses/by-sa/4.0/}{CC BY-SA 4.0}}
\date{Last updated: \today}

\begin{document}

\maketitle

\begin{itemize}[itemsep=3ex]

	\item Names of the review team members

	\item Names of the design team members

	\item Has the design team provided a bill of materials?

	\item Has the design team made sure that they will have all the parts needed before the PCB comes back? Please check every part in the circuit. 

	\item Has the design team provided datasheets, if available, to all the parts?

	\item Are all the polarized capacitors, if there are any, placed correctly?

	\item For all active parts, have the power and ground connections been properly placed?

	\item For all parts, does the circuit schematic follow suggestions from the datasheet? If not, what changes have been made and why?

	\item For all parts, does the PCB layout follow suggestions from the datasheet? If not, what changes have been made and why?

	\item Are the input and output (including power and signal) of the circuit properly arranged? Will there be difficulties in physically connecting to the input and output?
	
	\item Has the width of the PCB traces supplying current to active components been properly sized?
	 
	\item Have test points been included in critical positions in the circuit?

	\item Does the layout agree with the schematic? Please check every connection. 

	\item Are all layers of the PCB board properly poured with copper? Are the poured copper all connected to ground? If not, why?

	\item Have the output files, such as gerber files, been prepared?

\end{itemize}

\end{document}
